---
title: "Project 1 Markdown"
author: "Rachel Keller"
date: "September 18, 2020"
output: rmarkdown::github_document
---

# Requred Packages

Required packages for this vignette are `httr`, `jsonlite`, and `tidyverse`. We will now load in these packages. 
```{r, message = FALSE, warning = FALSE}
library(httr)
library(jsonlite)
library(tidyverse)
```

# Functions

In this section, we will create the functions that will contact the NHL Records API and NHL Stats API, and return parsed data for certain endpoints.

## Franchise Endpoint

The function below will return the franchise endpoint from the NHL Records API, given either the team name or team number/id.
```{r}
getFranchise <- function(teamName = NA, teamNumber = NA){
  baseURLNHLRecords <- "https://records.nhl.com/site/api"
  tableName <- "franchise"
  fullURL <- paste0(baseURLNHLRecords, "/", tableName)
  get <- GET(fullURL)
  txt <- content(get, "text", encoding = "UTF-8")
  json <- fromJSON(txt, flatten = TRUE)
  if(is.na(teamName) & is.na(teamNumber)){
    result <- json$data
  }
  else if (is.na(teamName)){
    result <- json$data %>% mutate(fullName = paste(teamPlaceName, teamCommonName, sep = " ")) %>% filter(id == teamNumber)
  }
  else {
    result <- json$data %>% mutate(fullName = paste(teamPlaceName, teamCommonName, sep = " ")) %>% filter(fullName == teamName)
  }
  return(result)
}
```

## Franchise Team Totals

The function below will return the franchise team totals endpoint from the NHL Records API, given either the team name or team number/id.
```{r}
getFranchiseTotals <- function(teamName = NA, teamNumber = NA){
  baseURLNHLRecords <- "https://records.nhl.com/site/api"
  tableName <- "franchise-team-totals "
  fullURL <- paste0(baseURLNHLRecords, "/", tableName)
  get <- GET(fullURL)
  txt <- content(get, "text", encoding = "UTF-8")
  json <- fromJSON(txt, flatten = TRUE)
  if(is.na(teamName) & is.na(teamNumber)){
    result <- json$data
  }
  else if (is.na(teamName)){
    result <- json$data %>% filter(franchiseId == teamNumber)
  }
  else {
    teamNameValue <- teamName
    result <- json$data %>% filter(teamName == teamNameValue)
  }
  return(result)
}
```

## Franchise Season Records

The function below will return the franchise season records endpoint from the NHL Records API, given either the team name or team number/id.
```{r}
getFranchiseSeasonRecords <- function(teamName = NA, teamNumber = NA){
  baseURLNHLRecords <- "https://records.nhl.com/site/api"
  tableName <- "franchise-season-records?cayenneExp=franchiseId="
  if(is.na(teamName) & is.na(teamNumber)){
    result <- "Error, need to specify a franchise ID or name. "
  }
  else if (is.na(teamName)){
    id <- teamNumber
    fullURL <- paste0(baseURLNHLRecords, "/", tableName, id)
    get <- GET(fullURL)
    txt <- content(get, "text", encoding = "UTF-8")
    json <- fromJSON(txt, flatten = TRUE)
    result <- json$data
  }
  else {
    teamInfo <- getFranchise(teamName = teamName)
    id <- teamInfo$id
    fullURL <- paste0(baseURLNHLRecords, "/", tableName, id)
    get <- GET(fullURL)
    txt <- content(get, "text", encoding = "UTF-8")
    json <- fromJSON(txt, flatten = TRUE)
    result <- json$data
  }
  return(result)
}
```

## Franchise Goalie Records

The function below will return the goalie records endpoint from the NHL Records API, given either the team name or team number/id.
```{r}
getGoalieRecords <- function(teamName = NA, teamNumber = NA){
  baseURLNHLRecords <- "https://records.nhl.com/site/api"
  tableName <- "franchise-goalie-records?cayenneExp=franchiseId="
  if(is.na(teamName) & is.na(teamNumber)){
    result <- "Error, need to specify a franchise ID or name. "
  }
  else if (is.na(teamName)){
    id <- teamNumber
    fullURL <- paste0(baseURLNHLRecords, "/", tableName, id)
    get <- GET(fullURL)
    txt <- content(get, "text", encoding = "UTF-8")
    json <- fromJSON(txt, flatten = TRUE)
    result <- json$data
  }
  else {
    teamInfo <- getFranchise(teamName = teamName)
    id <- teamInfo$id
    fullURL <- paste0(baseURLNHLRecords, "/", tableName, id)
    get <- GET(fullURL)
    txt <- content(get, "text", encoding = "UTF-8")
    json <- fromJSON(txt, flatten = TRUE)
    result <- json$data
  }
  return(result)
}
```

## Franchise Skater Records

The function below will return the skater records endpoint from the NHL Records API, given either the team name or team number/id.
```{r}
getSkaterRecords <- function(teamName = NA, teamNumber = NA){
  baseURLNHLRecords <- "https://records.nhl.com/site/api"
  tableName <- "franchise-skater-records?cayenneExp=franchiseId="
  if(is.na(teamName) & is.na(teamNumber)){
    result <- "Error, need to specify a franchise ID or name. "
  }
  else if (is.na(teamName)){
    id <- teamNumber
    fullURL <- paste0(baseURLNHLRecords, "/", tableName, id)
    get <- GET(fullURL)
    txt <- content(get, "text", encoding = "UTF-8")
    json <- fromJSON(txt, flatten = TRUE)
    result <- json$data
  }
  else {
    teamInfo <- getFranchise(teamName = teamName)
    id <- teamInfo$id
    fullURL <- paste0(baseURLNHLRecords, "/", tableName, id)
    get <- GET(fullURL)
    txt <- content(get, "text", encoding = "UTF-8")
    json <- fromJSON(txt, flatten = TRUE)
    result <- json$data
  }
  return(result)
}
```

## Teams

The function below will return the teams endpoint from the NHL Stats API, given either the team name or team number/id.
```{r}
getTeams <- function(teamName = NA, teamNumber = NA, modifier = NA, modifierValue = NA){
  baseURLNHLStats <- "https://statsapi.web.nhl.com/api/v1"
  tableName <- "teams"
  switch(modifier,
             "Team Roster"={
               modifierURL <- "?expand=team.roster"
             }, "Person Names"={
               modifierURL <- "?expand=person.names"
             }, "Next Schedule"={
               modifierURL <- "?expand=team.schedule.next"
             }, "Previous Schedule"={
               modifierURL <- "?expand=team.schedule.previous"
             }, "Team Stats"={
               modifierURL <- "?expand=team.stats"
             }, "Season Roster"={
               modifierURL <- paste0("?expand=team.roster&season=", modifierValue)
             }, "Multiple Teams"={
               modifierURL <- "?teamId="
             }, "Single Season Stats"={
               modifierURL <- "?stats=statsSingleSeasonPlayoffs"
             })
  if(is.na(teamName) & is.na(teamNumber)){
    fullURL <- paste0(baseURLNHLStats, "/", tableName)
    get <- GET(fullURL)
    txt <- content(get, "text")
    json <- fromJSON(txt, flatten=TRUE)
    result <- json$teams
  }
  else if (is.na(teamName)){
    fullURL <- paste0(baseURLNHLStats, "/", tableName)
    get <- GET(fullURL)
    txt <- content(get, "text")
    json <- fromJSON(txt, flatten=TRUE)
    if (is.na(modifier)){
      result <- json$teams %>% filter(franchiseId == teamNumber)
    } else {
      getId <- json$teams %>% filter(franchiseId == teamNumber)
      id <- getId$id
      URLWithModifier <- paste0(baseURLNHLStats, "/", tableName, "/", id, modifierURL)
      getModifier <- GET(URLWithModifier)
      text <- content(getModifier, "text")
      jsonFile <- fromJSON(text, flatten=TRUE)
      if (modifier == "Team Roster"){
        result <- jsonFile$teams$roster.roster
      } else if (modifier == "Previous Schedule"){
        result <- jsonFile$teams$previousGameSchedule.dates
      } else if (modifier == "Team Stats"){
        result <- jsonFile$teams$teamStats
      } else if (modifier == "Season Roster"){
        result <- jsonFile$teams$roster.roster
      }
      else {
        result <- jsonFile$teams
      }
    }
  }
  else {
    fullURL <- paste0(baseURLNHLStats, "/", tableName)
    get <- GET(fullURL)
    txt <- content(get, "text")
    json <- fromJSON(txt, flatten=TRUE)
    teamNameValue <- teamName
    if (is.na(modifier)){
      result <- json$teams %>% filter(name == teamNameValue)
    } else {
      getId <- json$teams %>% filter(name == teamNameValue)
      id <- getId$id
      URLWithModifier <- paste0(baseURLNHLStats, "/", tableName, "/", id, modifierURL)
      getModifier <- GET(URLWithModifier)
      text <- content(getModifier, "text")
      jsonFile <- fromJSON(text, flatten=TRUE)
      if (modifier == "Team Roster"){
        result <- jsonFile$teams$roster.roster
      } else if (modifier == "Previous Schedule"){
        result <- jsonFile$teams$previousGameSchedule.dates
      } else if (modifier == "Team Stats"){
        result <- jsonFile$teams$teamStats
      } else if (modifier == "Season Roster"){
        result <- jsonFile$teams$roster.roster
      }
      else {
        result <- jsonFile$teams
      }
    }
  }
  return(result)
}
```


## Wrapper

Now, we will create our wrapper function. The user can select their desired endpoint. 
```{r}
getEndpoint <- function(endpoint, teamName = NA, teamNumber = NA, modifier = NA, modifierValue = NA){
  if(endpoint == "franchise"){
    result <- getFranchise(teamName, teamNumber)
  } else if (endpoint == "franchise totals"){
    result <- getFranchiseTotals(teamName, teamNumber)
  } else if (endpoint == "franchise season records"){
    result <- getFranchiseSeasonRecords(teamName, teamNumber)
  } else if (endpoint == "goalie records"){
    result <- getGoalieRecords(teamName, teamNumber)
  } else if (endpoint == "skater records"){
    result <- getSkaterRecords(teamName, teamNumber)
  } else if (endpoint == "teams"){
    result <- getTeams(teamName, teamNumber, modifier, modifierValue)
  } else {
    result <- "Error, incorrect endpoint."
  }
  return(result)
}
```

# Exploratory Data Analysis

Now that we have all of functions created to contact various endpoints in our APIs, we will begin exploring our data. 

We'll begin by doing some analysis for the NHL overall, then dive into specific details the Washington Capitals. Examin some metrics that could indicate good team performance. 

More likely to win a shootout
Franchise totals data
Create boxplot
create new variable
```{r, warning = FALSE}
shootout <- getEndpoint("franchise totals") %>% mutate(propShootouts = shootoutWins / (shootoutWins + shootoutLosses), na.rm=TRUE)
g <- ggplot(shootout, aes(X = teamName, y = propShootouts, fill = teamName))
g + geom_boxplot()
```


Bar graph of home wins by division
```{r}
teamInfo <- getEndpoint("teams")
joined <- right_join(teamInfo, shootout, by = "franchiseId")
g <- ggplot(joined, aes(name, homeWins))
g + geom_bar(stat = "identity") + facet_grid(~division.name) + coord_flip()
```

Road wins by division
```{r}
teamInfo <- getEndpoint("teams")
joined <- right_join(teamInfo, shootout, by = "franchiseId")
g <- ggplot(joined, aes(name, roadWins))
g + geom_bar(stat = "identity") + facet_grid(~division.name) + coord_flip()
```

```{r}
joined %>% filter(gameTypeId == 2) %>% group_by(name) %>% summarise(avgPoints = mean(points))
joined %>% filter(gameTypeId == 2) %>% group_by(division.name) %>% summarise(avgPoints = mean(points))
```


average points per game in regular season
```{r}
joined %>% filter(gameTypeId == 2) %>% mutate(avgPoints = points / gamesPlayed, na.rm=TRUE) %>% group_by(division.name) %>% summarise(avg = mean(avgPoints), med = median(avgPoints), var = var(avgPoints))
joined %>% filter(gameTypeId == 2) %>% mutate(avgPoints = points / gamesPlayed, na.rm=TRUE) %>% group_by(name) %>% summarise(avg = mean(avgPoints))
```

Use skater records for capitals and make histogram of seasons
Skater records data
Histogram
```{r}
seasons <- getEndpoint("skater records", teamName = "Washington Capitals")
g <- ggplot(seasons, aes(x = seasons))
g + geom_histogram(binwidth = 3)
```


Use skater records for capitals and make histogram of seasons
Goalie records data
Histogram
```{r}
seasons <- getEndpoint("goalie records", teamName = "Washington Capitals")
g <- ggplot(seasons, aes(x = seasons))
g + geom_histogram(binwidth = 3)
```


```{r}
seasons <- getEndpoint("skater records", teamName = "Washington Capitals")
g <- ggplot(seasons, aes(x = mostPointsOneSeason, y = mostPointsOneGame))
g + geom_point()+ facet_wrap(~activePlayer)
```


Make scatterplot of with franchise season records of win streak vs loss streak

```{r}
seasons <- getEndpoint("goalie records", teamName = "Washington Capitals")
g <- ggplot(seasons, aes(x = mostShotsAgainstOneGame, y = mostSavesOneGame))
g + geom_point() + facet_wrap(~activePlayer)
```


contingency table of active players
```{r}
seasons <- getEndpoint("skater records", teamName = "Washington Capitals")
table(seasons$activePlayer)
```

contingency table of active goalies
```{r}
seasons <- getEndpoint("goalie records", teamName = "Washington Capitals")
table(seasons$activePlayer)
```
```{r}
positions <- getEndpoint("teams", teamName = "Washington Capitals", modifier = "Team Roster")
table(positions[[1]][["position.name"]])
```

```{r}
seasons <- getEndpoint("skater records", teamName = "Washington Capitals")
g <- ggplot(seasons, aes(x = seasons, y = points, group = seasons))
g + geom_boxplot()
```

```{r}
seasons <- getEndpoint("skater records", teamName = "Washington Capitals")
g <- ggplot(seasons, aes(x = seasons, y = penaltyMinutes, group = seasons))
g + geom_boxplot()
```

