---
title: "Project 1 Vignette"
author: "Rachel Keller"
date: "September 18, 2020"
output: rmarkdown::github_document

---

# Table of Contents 

* Required Packages
* Functions
  + Franchise Endpoint
  + Franchise Team Totals
  + Franchise Season Records
  + Franchise Goalie Records
  + Franchise Skater Records
  + Teams
  + Wrapper
* Exploratory Data Analysis
  + NHL Analysis
  + Washington Capitals Analysis

# Requred Packages

The required packages for this vignette are `httr`, `jsonlite`, `tidyverse`, and `knitr`. We will load in these packages. 
```{r, message = FALSE, warning = FALSE}
library(httr)
library(jsonlite)
library(tidyverse)
library(knitr)
```

# Functions

In this section, we will create the functions that will contact the NHL Records API and NHL Stats API, and return parsed data for certain endpoints.

## Franchise Endpoint

The function below will return the franchise endpoint from the NHL Records API, given either the team name or team number/id. Variables returned include id, firstSeasonId and lastSeasonId and name of every team in the history of the NHL.
```{r}
getFranchise <- function(teamName = NA, teamNumber = NA){ 
  # Generate URL to contact API
  baseURLNHLRecords <- "https://records.nhl.com/site/api" 
  tableName <- "franchise"
  fullURL <- paste0(baseURLNHLRecords, "/", tableName)
  # Create get request
  get <- GET(fullURL)
  txt <- content(get, "text", encoding = "UTF-8")
  json <- fromJSON(txt, flatten = TRUE)
  if(is.na(teamName) & is.na(teamNumber)){
    # If no specified team return full data set
    result <- json$data
  }
  else if (is.na(teamName)){
    # If given teamNumber return filtered data set
    result <- json$data %>% mutate(fullName = paste(teamPlaceName, teamCommonName, sep = " ")) %>% filter(id == teamNumber)
  }
  else {
    # If given teamName return filtered data set
    result <- json$data %>% mutate(fullName = paste(teamPlaceName, teamCommonName, sep = " ")) %>% filter(fullName == teamName)
  }
  return(result)
}
```

## Franchise Team Totals

The function below will return the franchise team totals endpoint from the NHL Records API, given either the team name or team number/id. Information returned from this endpoint include total stats for every franchise such as roadTies, roadWins, etc. 
```{r}
getFranchiseTotals <- function(teamName = NA, teamNumber = NA){
  # Generate URL to contact API
  baseURLNHLRecords <- "https://records.nhl.com/site/api"
  tableName <- "franchise-team-totals "
  fullURL <- paste0(baseURLNHLRecords, "/", tableName)
  # Create get request
  get <- GET(fullURL)
  txt <- content(get, "text", encoding = "UTF-8")
  json <- fromJSON(txt, flatten = TRUE)
  if(is.na(teamName) & is.na(teamNumber)){
    # If no specified team return full data set
    result <- json$data
  }
  else if (is.na(teamName)){
    # If given teamNumber return filtered data set
    result <- json$data %>% filter(franchiseId == teamNumber)
  }
  else {
    # If given teamName return filtered data set
    teamNameValue <- teamName
    result <- json$data %>% filter(teamName == teamNameValue)
  }
  return(result)
}
```

## Franchise Season Records

The function below will return the franchise season records endpoint from the NHL Records API, given either the team name or team number/id. This endpoint drills down into season records for a specific franchise, so a team must be specified for this endpoint. 
```{r}
getFranchiseSeasonRecords <- function(teamName = NA, teamNumber = NA){
  baseURLNHLRecords <- "https://records.nhl.com/site/api"
  tableName <- "franchise-season-records?cayenneExp=franchiseId="
  if(is.na(teamName) & is.na(teamNumber)){
    # Endpoint needs a franchise to deliver results
    result <- "Error, need to specify a franchise ID or name. "
  }
  else if (is.na(teamName)){
    # Create URL to contact API
    id <- teamNumber
    fullURL <- paste0(baseURLNHLRecords, "/", tableName, id)
    # Create get request
    get <- GET(fullURL)
    txt <- content(get, "text", encoding = "UTF-8")
    json <- fromJSON(txt, flatten = TRUE)
    result <- json$data
  }
  else {
    # Look up id using getFranchise
    teamInfo <- getFranchise(teamName = teamName)
    id <- teamInfo$id
    # Create URL to contact API
    fullURL <- paste0(baseURLNHLRecords, "/", tableName, id)
    # Create get request
    get <- GET(fullURL)
    txt <- content(get, "text", encoding = "UTF-8")
    json <- fromJSON(txt, flatten = TRUE)
    result <- json$data
  }
  return(result)
}
```

## Franchise Goalie Records

The function below will return the goalie records endpoint from the NHL Records API, given either the team name or team number/id. This endpoint returns goalie records for the specified franchise.
```{r}
getGoalieRecords <- function(teamName = NA, teamNumber = NA){
  baseURLNHLRecords <- "https://records.nhl.com/site/api"
  tableName <- "franchise-goalie-records?cayenneExp=franchiseId="
  if(is.na(teamName) & is.na(teamNumber)){
    # Endpoint needs a franchise to deliver results
    result <- "Error, need to specify a franchise ID or name. "
  }
  else if (is.na(teamName)){
    # Create URL to contact API
    id <- teamNumber
    fullURL <- paste0(baseURLNHLRecords, "/", tableName, id)
    # Create get request
    get <- GET(fullURL)
    txt <- content(get, "text", encoding = "UTF-8")
    json <- fromJSON(txt, flatten = TRUE)
    result <- json$data
  }
  else {
    # Look up id using getFranchise
    teamInfo <- getFranchise(teamName = teamName)
    id <- teamInfo$id
    # Create URL to contact API
    fullURL <- paste0(baseURLNHLRecords, "/", tableName, id)
    # Create get request
    get <- GET(fullURL)
    txt <- content(get, "text", encoding = "UTF-8")
    json <- fromJSON(txt, flatten = TRUE)
    result <- json$data
  }
  return(result)
}
```

## Franchise Skater Records

The function below will return the skater records endpoint from the NHL Records API, given either the team name or team number/id. This endpoint returns skater records for the specified franchise.
```{r}
getSkaterRecords <- function(teamName = NA, teamNumber = NA){
  # Endpoint needs a franchise to deliver results
  baseURLNHLRecords <- "https://records.nhl.com/site/api"
  tableName <- "franchise-skater-records?cayenneExp=franchiseId="
  if(is.na(teamName) & is.na(teamNumber)){
    result <- "Error, need to specify a franchise ID or name. "
  }
  else if (is.na(teamName)){
    # Create URL to contact API
    id <- teamNumber
    fullURL <- paste0(baseURLNHLRecords, "/", tableName, id)
    # Create get request
    get <- GET(fullURL)
    txt <- content(get, "text", encoding = "UTF-8")
    json <- fromJSON(txt, flatten = TRUE)
    result <- json$data
  }
  else {
    # Look up id using getFranchise
    teamInfo <- getFranchise(teamName = teamName)
    id <- teamInfo$id
    # Create URL to contact API
    fullURL <- paste0(baseURLNHLRecords, "/", tableName, id)
    # Create get request
    get <- GET(fullURL)
    txt <- content(get, "text", encoding = "UTF-8")
    json <- fromJSON(txt, flatten = TRUE)
    result <- json$data
  }
  return(result)
}
```

## Teams

The function below will return the teams endpoint from the NHL Stats API, given either the team name or team number/id. The endpoint includes arguments for eight modifiers.
```{r}
getTeams <- function(teamName = NA, teamNumber = NA, modifier = NA, modifierValue = NA){
  baseURLNHLStats <- "https://statsapi.web.nhl.com/api/v1"
  tableName <- "teams"
  # Create URL given modifier input
  switch(modifier,
             "Team Roster"={
               modifierURL <- "?expand=team.roster"
             }, "Person Names"={
               modifierURL <- "?expand=person.names"
             }, "Next Schedule"={
               modifierURL <- "?expand=team.schedule.next"
             }, "Previous Schedule"={
               modifierURL <- "?expand=team.schedule.previous"
             }, "Team Stats"={
               modifierURL <- "?expand=team.stats"
             }, "Season Roster"={
               modifierURL <- paste0("?expand=team.roster&season=", modifierValue)
             }, "Multiple Teams"={
               modifierURL <- "?teamId="
             }, "Single Season Stats"={
               modifierURL <- "?stats=statsSingleSeasonPlayoffs"
             })
  if(is.na(teamName) & is.na(teamNumber)){
    fullURL <- paste0(baseURLNHLStats, "/", tableName)
    # Create get request and return full data set
    get <- GET(fullURL)
    txt <- content(get, "text")
    json <- fromJSON(txt, flatten=TRUE)
    result <- json$teams
  }
  else if (is.na(teamName)){
    # Contact full data set
    fullURL <- paste0(baseURLNHLStats, "/", tableName)
    get <- GET(fullURL)
    txt <- content(get, "text")
    json <- fromJSON(txt, flatten=TRUE)
    if (is.na(modifier)){
      # If no modifier return filtered version of full data set
      result <- json$teams %>% filter(franchiseId == teamNumber)
    } else {
      # Get id from full data set
      getId <- json$teams %>% filter(franchiseId == teamNumber)
      id <- getId$id
      URLWithModifier <- paste0(baseURLNHLStats, "/", tableName, "/", id, modifierURL)
      # Create get request
      getModifier <- GET(URLWithModifier)
      text <- content(getModifier, "text")
      jsonFile <- fromJSON(text, flatten=TRUE)
      # Return output depending on modifier value
      if (modifier == "Team Roster"){
        result <- jsonFile$teams$roster.roster
      } else if (modifier == "Previous Schedule"){
        result <- jsonFile$teams$previousGameSchedule.dates
      } else if (modifier == "Team Stats"){
        result <- jsonFile$teams$teamStats
      } else if (modifier == "Season Roster"){
        result <- jsonFile$teams$roster.roster
      }
      else {
        # Return full data set
        result <- jsonFile$teams
      }
    }
  }
  else {
    # Contact full data set
    fullURL <- paste0(baseURLNHLStats, "/", tableName)
    get <- GET(fullURL)
    txt <- content(get, "text")
    json <- fromJSON(txt, flatten=TRUE)
    teamNameValue <- teamName
    if (is.na(modifier)){
      # If no modifier return filtered version of full data set
      result <- json$teams %>% filter(name == teamNameValue)
    } else {
      # Get id from full data set
      getId <- json$teams %>% filter(name == teamNameValue)
      id <- getId$id
      URLWithModifier <- paste0(baseURLNHLStats, "/", tableName, "/", id, modifierURL)
      # Create get request
      getModifier <- GET(URLWithModifier)
      text <- content(getModifier, "text")
      jsonFile <- fromJSON(text, flatten=TRUE)
      # Return output depending on modifier value
      if (modifier == "Team Roster"){
        result <- jsonFile$teams$roster.roster
      } else if (modifier == "Previous Schedule"){
        result <- jsonFile$teams$previousGameSchedule.dates
      } else if (modifier == "Team Stats"){
        result <- jsonFile$teams$teamStats
      } else if (modifier == "Season Roster"){
        result <- jsonFile$teams$roster.roster
      }
      else {
        # Return full data set
        result <- jsonFile$teams
      }
    }
  }
  return(result)
}
```


## Wrapper

Now, we will create our wrapper function. The user can select their desired endpoint. 
```{r}
getEndpoint <- function(endpoint, teamName = NA, teamNumber = NA, modifier = NA, modifierValue = NA){
  # Select function depending on endpoint desired
  if(endpoint == "franchise"){
    result <- getFranchise(teamName, teamNumber)
  } else if (endpoint == "franchise totals"){
    result <- getFranchiseTotals(teamName, teamNumber)
  } else if (endpoint == "franchise season records"){
    result <- getFranchiseSeasonRecords(teamName, teamNumber)
  } else if (endpoint == "goalie records"){
    result <- getGoalieRecords(teamName, teamNumber)
  } else if (endpoint == "skater records"){
    result <- getSkaterRecords(teamName, teamNumber)
  } else if (endpoint == "teams"){
    result <- getTeams(teamName, teamNumber, modifier, modifierValue)
  } else {
    # If no or incorrect endpoint specified
    result <- "Error, incorrect endpoint."
  }
  return(result)
}
```

# Exploratory Data Analysis

Now that we have all of functions created to contact various endpoints in our APIs, we will explore our data. We'll begin by doing some analysis for the overall NHL, then dive into analysis of the Washington Capitals, my personal team. For both analyses, we will take a look at a few measures that could indicate strong performance. 

## NHL Analysis

First, we will look at a contingency table to get an idea about the overall structure of the NHL.
```{r}
overallStructure <- getEndpoint("teams")
kable(table(overallStructure$division.name, overallStructure$conference.name))
```

We can see that there are 16 active teams in the Eastern conceference, 8 in the Atlantic division and 8 in the Metropolitan division. And there are 15 active teams in the Western conference, 7 in the Central division and 8 in the Pacific division. 

As standings are determined by points, we will create numeric summaries with the overall points by team, division, and conference.
```{r, message=FALSE}
conferences <- getEndpoint("teams") %>% rename(teamId = id)
points <- getEndpoint("franchise totals") %>% filter(gameTypeId == 2)
conferencesPoints <- right_join(conferences, points, by = "teamId")
conferencesPoints %>% group_by(name) %>% summarise(pointsEarned = points) %>% filter(!is.na(name))
conferencesPoints %>% group_by(division.name) %>% summarise(sumPointsEarned = sum(points)) %>% filter(!is.na(division.name))
conferencesPoints %>% group_by(conference.name) %>% summarise(sumPointsEarned = sum(points)) %>% filter(!is.na(conference.name))
```
We can see that the Montréal Canadiens have the most amount of points earned over the entire history of the NHL at 7899. Also, the Atlantic division and Eastern conferences have the largest amount of points. The Canadiens belong to the Eastern conference and Atlantic division, so this partly explains why that division and conference have the largest amount of points. 

Now that we have looked at points as a whole, we will look at the average points per game in regular season by team, division, and conference.
```{r, message=FALSE}
conferencesPoints %>% mutate(avgPoints = points / gamesPlayed) %>% group_by(name) %>% summarise(avg = mean(avgPoints)) %>% filter(!is.na(name))
conferencesPoints %>% mutate(avgPoints = points / gamesPlayed) %>% filter(!is.na(name)) %>% group_by(division.name) %>% summarise(avg = mean(avgPoints), med = median(avgPoints), var = var(avgPoints))
conferencesPoints %>% mutate(avgPoints = points / gamesPlayed) %>% filter(!is.na(name)) %>% group_by(conference.name) %>% summarise(avg = mean(avgPoints), med = median(avgPoints), var = var(avgPoints))
```
The Vegas Golden Knights have the highest average of points per game. Also, the Cental division and Western conference have the highest per game point average. We can see that median of the Cental division and median of the Western conference is greater than the average, which suggests that the data is skewed left. The Pacific division has a noticably higher spread. And the Western conference also has a higher spread than the Eastern conference. 

Now that we have done some analysis on points, we will begin to look at the some metrics that may imdicate good performance, such as the number of wins overall. First, we will create a bar graph of home wins by conference.
```{r}
teamInfo <- getEndpoint("teams") %>% rename(teamId = id)
totals <- getEndpoint("franchise totals")
teamInfoTotals <- right_join(teamInfo, totals, by = "teamId") %>% filter(!is.na(name), !is.na(conference.name))
g <- ggplot(teamInfoTotals, aes(name, homeWins, fill = conference.name))
g + geom_bar(stat = "identity") + facet_grid(~conference.name) + coord_flip() + labs(y = "Number of Wins at Home", x = "Team Name", title = "Bar Plot of Home Wins by Conference", fill = "Conference")
```
In the Eastern conference, the Montréal Canadiens have the largest number of wins at home. And in the Western conference, the Chicago Blackhawks have the largest number of home wins. 

Staying on the topic of wins, we will now look at a bar plot of road wins by conference. 
```{r}
g <- ggplot(teamInfoTotals, aes(name, roadWins, fill = conference.name))
g + geom_bar(stat = "identity") + facet_grid(~conference.name) + coord_flip() + labs(y = "Number of Wins on the Road", x = "Team Name", title = "Bar Plot of Road Wins by Conference", fill = "Conference")
```
Not surprisingly, the Montréal Canadiens have the largest number of wins on the road for the Eastern conference. Also, the Chicago Blackhawks have the largest number of road wins for the Western conference. 

The last performance metric we will analyze are shootouts. We will take a look at the distribution of the proportion of shootouts won, for both regular and post season. 
```{r, warning = FALSE}
shootout <- getEndpoint("franchise totals") %>% mutate(propShootouts = shootoutWins / (shootoutWins + shootoutLosses))
g <- ggplot(shootout, aes(X = teamName, y = propShootouts, fill = teamName))
g + geom_boxplot() + labs(x = "Team", y = "Proportion of Successful Shootouts", title = "Bar Plot of Shootouts")
```
We can see that the vast majority of teams hover around the 0.5 mark. However, 4 teams have distributions that span from 0.5 to 1, and from 0 to 0.5. These teams, the Washington Capitals, Tampa Bay Lighting, St. Louis Blues, and Dallas Stars have either won or lost all of their shootouts in the post season. 

## Washington Capitals Analysis

Much like the analysis of the NHL, we will begin by creating some contingency tables to gain some insight as to the layout of the Capitals.

First, we will create a contingency table of the skater history. 
```{r}
skaterHistory <- getEndpoint("skater records", teamName = "Washington Capitals")
kable(table(Active = skaterHistory$activePlayer))
```

There are currently 52 active skaters for the Washington Capitals. 

And now we will create a similar contingency table of the goalie history. 
```{r}
goalieHistory <- getEndpoint("goalie records", teamName = "Washington Capitals")
kable(table(Active = goalieHistory$activePlayer))
```

There are currently 3 active goalies on the team. 

We will create another contingency table that further drills down into the positions of the current roster. 
```{r}
currentRoster <- getEndpoint("teams", teamName = "Washington Capitals", modifier = "Team Roster")
kable(table(Position = currentRoster[[1]][["position.name"]]))
```

The position with the most players is defensemen, followed by center and left wing. 

Now we will start to dig into the details of the players, first by seeing how much experience they have. To do this, we will create a histogram of the seasons a player has been with the Capitals, for both skaters and goalies. 
```{r, message = FALSE}
skaterSeasons <- getEndpoint("skater records", teamName = "Washington Capitals") %>% mutate(skater = "Yes")
goalieSeasons <- getEndpoint("goalie records", teamName = "Washington Capitals") %>% mutate(skater = "No")
joinedSeasons <- full_join(skaterSeasons, goalieSeasons)
g <- ggplot(joinedSeasons, aes(x = seasons, fill = skater))
g + geom_histogram(binwidth = 3) + facet_wrap(~skater) + labs(x = "Number of Seasons", y = "Number of Players", title = "Distribution of Seasons for Players", fill = "Skater")
```

Both goalie and skater distributions appear to be right skewed. There are very few players who have a long career with the Capitals. 

Players with more experience should also perform better. We will test this by looking at boxplots of points by season earned by each skater and goalie.  
```{r, warning = FALSE}
g <- ggplot(joinedSeasons, aes(x = seasons, y = points, group = seasons))
g + geom_boxplot(color = "blue") + labs(x = "Number of Seasons", y = "Points Earned", title = "Distribution of Points by Seasons")
```

There is a steady upward trend. As players gain more experience and play for more seasons, the points they earn are consistently higher. Or alternatively, players who no longer earn higher points are drafted to another team and/or no longer play for the NHL. 

Finally, we will look at the points earned for skaters, both active and inactive, in one season vs. one game, to see if the Capitals are improving in their choice of skaters over time. 
```{r}
seasonPoints <- getEndpoint("skater records", teamName = "Washington Capitals")
g <- ggplot(seasonPoints, aes(x = mostPointsOneSeason, y = mostPointsOneGame, color = activePlayer))
g + geom_jitter() + labs(x = "Most Points Earned in One Season", y = "Most Points in One Game", title = "Points Earned in One Season and One Game", color = "Active Player")
```

The trend of both active and nonactive players appears to be very similar. So the Capitals are not necessarily improving their choice of skaters over time. 

Now that we have looked at skaters, we will begin looking at some stats that may represent a solid goalie. We will look at a scatterplot of shots against one vs. the saves in one game, for both active and inactive goalies. 
```{r}
seasons <- getEndpoint("goalie records", teamName = "Washington Capitals")
g <- ggplot(seasons, aes(x = mostShotsAgainstOneGame, y = mostSavesOneGame, color = activePlayer))
g + geom_jitter() + labs(x = "Most Shots Against One", y = "Most Saves in One Game", title = "Shots Against One and Save in a Game", color = "Active Player")
```

There is a general upward trend in the scatterplot. The current goalie for the Capitals seems to fall right in line with goalies of the past. 

The last statistics we will take a look at are penalty minutes. To do this, we will create a boxplot of penalty minutes by the number of seasons a player has played for the Washington Capitals. .
```{r, warning = FALSE}
g <- ggplot(joinedSeasons, aes(x = seasons, y = penaltyMinutes, group = seasons))
g + geom_boxplot(color = "blue") + labs(x = "Number of Seasons", y = "Penalty Minutes", title = "Distribution of Penalties by Seasons")
```

There is a clear trend here, the longer a player has stuck around, the the longer they will end up sitting in the penalty box. 